### 2017-05-01 paper/samplebody-journals.tex
images

\quad Our first objective is to determine if certain attributes are indicative of business success with respect to their average review rating (see sec. 2.3 for revenue relation). For the businesses with many reviews, time-series data of the average star rating of the business mined from user reviews can be analyzed with ML models like Hidden Markov Models or recurrent neural networks in order to learn the likelihood of a business failing or succeeding in the future based on its current ratings. Furthermore, sentiment analysis can be used on language of reviews and matched with the star reviews to provide more information on a business's chances of success. Additionally, information about the current state of a business and likely future states could be provided to businesses to permit them to change their current practices and focus to increase their chances of success.

@@

A previous winner, Hood et al., published a paper on assessing a business's current rating/review state and future expectations with respect to review numbers on Yelp. The bulk of their work centered around the creation of additional features for the data (e.g. number of similar businesses within 1km, features of subsets of reviews such as average rating and number of unique users). From there they utilized various feature reduction and clustering techniques to build a prediction model. They were able to identify features that offered greater accuracy in prediction of interest in the business (calculated as the predicted number of reviews within 6 months after a target date) as well as a create model for prediction that outperformed a basic linear regression.

@@

Huang et al. (2013) utilized probabilistic models to discover under- lying subtopics within Yelp reviews. This offered a framework to categorize the reviews based on review text. From here, they were able to analyze the prevalence of various star ratings within each category. In this way, they could advise business owners as to the common trends among good and bad reviews that may not be apparent without considering the review text. McAuley and Leskovec built on this idea by combining such latent review topics with hidden rating driving factors, allowing them construct a rating prediction model.

@@

\quad The dataset has many optional attributes. There is an attribute simply called 'attributes' that is a shopping list of optional fields for each business. This part of the dataset is inconsistent and the same fields cannot be relied upon to be present for each business. For instance, one of these attributes is 'Bike Parking.' This attribute is either true, false, or nonexistent. In the cases when it is nonexistent, analysis based upon this attribute is obviously unreliable. Most of these miscellaneous attributes are present for much less than a majority of the data, so any mining based upon them would have to first drastically trim the data available for analysis by eliminating more than half the data. For this reason, we have chosen to eliminate the miscellaneous category rather than most of our data.

\quad In addition, the data is not only restaurants but many kinds of businesses. A datum associated with each business attempts to classify the business as a member of a category, such as 'restaurants,' 'beaches,' or 'shopping.' However some categories are extremely broad and the same type of business can be described in different ways. One example is that a bakery may be classified as belonging to 'bakeries,' but also generically as 'food.' Conversely, many businesses sell food, but the generic category 'food' will not be in their list of categories and instead the business will be labelled with more specific categories like 'Italian food,' diminishing the usefulness of these categories.

\quad The dataset provided by the Yelp organizers is fairly clean already, and the majority of the data cleaning process will be involved in handling missing data. It is possible to fill in these missing values with default values like 'false', or 'none' but it can be dangerous to make assumptions about what businesses have and don't have - for example, filling in a 'gluten-free' attribute of a business may influence our analysis of its reviews and, depending on exposure, maybe even mislead Yelp users as to what the business provides.

@@

\quad The Yelp dataset comes from the Yelp Dataset Challenge. This is an ongoing challenge and this is the ninth round. The challenge offers \$5,000 to the winner and the  final submission is an academic paper.

@@

\begin{itemize}
	\item{4.1 million reviews}
	\begin{verbatim}
	"review_id":(encrypted review id),
	"user_id":(encrypted user id),
	"business_id":(encrypted business id)"
	"stars":(star rating, rounded to half-stars),
	"date":(date, formatted: YYYY-MM-DD),
	"text":"review text",
	"useful":('useful' vote count),
	"funny":('funny' vote count),
	"cool": ('cool' vote count),
	"type": "review"
	\end{verbatim}
	\item{Information on 144,000 businesses in 11 cities from 4 countries (incl. the US, Germany, Canada, and the U.K. (with 1.1 million business attributes)}
	\begin{verbatim}
	"business_id":(encrypted business id),
	"name":"business name",
	"neighborhood": "hood name",
	"address":"full address",
	"city":"city",
	"state":"state"(if applicable),
	"postal code":"postal code",
	"latitude":latitude,
	"longitude":longitude,
	"stars":(star rating, rounded to half-stars),
	"review_count":(number of reviews),
	"is_open":0/1 (closed/open),
	"attributes":[(localized attribute tags)],
	"categories":[(localized category names)],
	"hours":[(hours strings)],
	"type": "business"
	\end{verbatim}
	\item{Over 1M users}
	\begin{verbatim}
	"user_id":(encrypted user id),
	"name":"first name",
	"review_count":(number of reviews),
	"yelping_since": (date, formatted: YYYY-MM-DD),
	"friends":[(encrypted user ids)],
	"useful":('useful' vote count sent by user),
	"funny":('funny' vote count sent by user),
	"cool": ('cool' vote count sent by user),
	"fans":"number of fans the user has",
	"elite":["an array of years the user was elite"],
	"average_stars":floating point average like 4.31,
	"compliment_type": (compliment count),
		...(same for each compliment type)...
	"type":"user"
	\end{verbatim}
	\item{Over 200,000 user-submitted photographs, 947,000 tips, and aggregated check-in data}
\end{itemize}

@@

\quad The tools and methods used for this project will no longer include a database for storing or integrating the data. Due to difficulties described in section 6 and relating to the unknown JSON variant used by the preparers of the Yelp dataset, storing the data in a database is not feasible at this time. Given that the data is already quite clean save for the format, with minimal replacement of missing values and some other changes necessary, it is possible for each group member to have a copy of the data stored locally that may be processed with our tools, with smaller subsets of data, results, and scripts shared via GitHub.

\quad The datasets will be merged locally, and while this does increase the amount of computational work necessary for each member?' machine, the dataset is still of a small enough size (less than 10 GB, excluding the photos) that this is practical. We will still use various Python libraries and modules for cleaning and preprocessing, and various other tools for transformation, mining, pattern evaluation, and knowledge presentation. The lack of a centralized database in removes the need for a number of Python libraries like Dora and FTFY ('fixes text for you') intended for data cleaning and preprocessing, but increases the necessity of group communication to reduce duplication of work and miscommunications. We currently use and will continue to use Slack, email, and in-person meetings to communicate and synchronize our efforts.

\quad Numpy, Scipy, and Pandas will be used to transform and mine the data, and other Python modules like Seaborn and Plotly (a Python API for D3) will be used for knowledge visualization and presentation. Rather than relying on tools like Weka, the Text Analytics API from Microsoft Cognitive Services, or the Facebook fastText pre-trained word vectors, we will instead use Python's Natural Language Toolkit for sentiment analysis and other language processing.

@@

\quad Evaluating our predictions as to whether a business success will hinge on comparing the star rating (from one to  ve stars) and our predicted star rating based on the other attributes in the review, such as day of the week and month, season, or location. This could be done by training an ML algorithm like Kernel Ridge regression on 80\% of the data and testing it on the remaining 20\%, and then adjusting parameters to minimize the error between the actual star ratings and our predicted star ratings for the testing data.

@@

\quad There have been several major changes to the project tools and methods since the project proposal. Initially we planned on using a relational database to store, integrate, and access the data, but but we encountered many difficulties in just uploading the data. The inconsistencies between the user and business data made it exceedingly difficult to successfully import the data en masse into the database, so instead we decided to import the data straight into Weka. However, the format of the JSON file also did not allow a successful import into Weka.

@@

\quad After importing the CSV formatted version of the data we tested some basic visualization of the data. For instance, in Figure 1 we created a histogram of the average star for a business versus the state in which it's located. Basic graphs like these help refine our question. For example, why does Illinois have significantly lower average stars than others?

\quad The most pervasive issue with our business data is that many entries are missing data, most notably the 'attributes' and address information. Since these values are managed by the account owner, they vary widely across the businesses.

\begin{figure}[h]
\includegraphics[width=0.5\textwidth, center]{b_rating_by_state}
\label{fig:Average star rating by state}
\caption{Note the outliers IL and NTH}
\end{figure}

Figure 2, depicts a graph of longitude vs. latitude vs. review count for businesses.

\begin{figure}[!h]
\includegraphics[width=0.5\textwidth, center]{3d_businesses_lat_long}
\caption{Clusters likely indicate areas of high population}
\label{fig:Review count by latitude and longitude}
\end{figure}

\quad In this instance the visualization makes it clear that reviews are more frequent in populated cities as we expect. However, it also shows visually the outliers. There are various businesses that have only a few reviews. In this context it is easier to identify outliers and also see that businesses concentrate in certain locations. Therefore we can reduce the dataset to only the cities that meet the threshold of businesses or total review counts.

@@

\quad One of the major goals of the project is to find recurrent problems and other issues with the businesses, and this was to be done by grouping reviews by business and then analyzing trends in the language used in the reviews over time. However, given that we were unable to find a suitable workaround for our difficulties with Weka, we used sentiment analysis tools from the Natural Language Toolkit (NLTK) to analyze the review language instead. NLTK is a language-processing and analysis platform that uses Python, and a good replacement for the language processing capabilities of Weka.

\quad Using a custom script (see GitHub; nltk\_review\_subset\_process.py) to analyze the sentiment of the first 100,000 reviews (about 2.5\% of the available data), we discovered that the average sentiment score among all analyzed reviews was around 0.113, and there were 17295 negative reviews compared to 76968 positive reviews. The remaining 5737 reviews had composite sentiment scores within -0.2 < score < 0.2, so they were excluded from further analysis since they were neutral. This means that the average Yelp review (at least within the analyzed subset) is positive rather than negative or neutral in sentiment, perhaps indicating that Yelp users are more inclined to write reviews of businesses with which they had a good experience. Only around 2.5\% of the data is currently processed by the file, because processing the entire file is prohibitively expensive.

@@

\quad There are some very interesting features here. It essentially provides a window into users' status, connections, and interaction within the Yelp ecosystem. The figure below shows a plot of the user data. We see that the user density grows around 3.25-4.25 average rating with some traces of a bell curve regarding the number of reviews. In face, it appears that the user data skews so heavily in this grouping that all other measures

\begin{figure}[!h]
\includegraphics[width=0.5\textwidth, center]{u_reviews_vs_stars}
  \caption{Note how the density increases between 3.25 and 4.25 starts}
  \label{fig:Plot of review count vs average star rating}
\end{figure}

\quad One of the most significant yet hardest to process attributes is the 'friends' list. This attribute provides an opportunity to model our data as a graph. Such a network could provide very relevant clustering through sparseness measures and cutting in addition to accounting for user/business influence or finding deeper trends according to network connectivity.  
\quad
\section{Next Steps}

@@

\quad Currently, the review dataset is too large to process in detail. In the future, we will continue taking subset samples of a more reasonable size and processing those instead, as well as increasing the complexity of our analysis. Combining the review dataset with the business dataset (using the ?business\_id? field as a key) will enable us to produce maps of where the largest number of reviews with negative and positive sentiment are, as well as seeing where users leave the most strongly-worded reviews. We hope that additional advanced sentiment analysis will be useful in determining underlying factors that influence a business's success, as well as in identifying recurrent problems.

>>>>

\quad Our first objective is to determine if certain attributes are indicative of business success with respect to their average review rating (see sec. 2.3 for revenue relation). For the businesses with many reviews, time-series data of the average star rating of the business mined from user reviews can be analyzed with ML models like Hidden Markov Models or recurrent neural networks in order to learn the likelihood of a business failing or succeeding in the future based on its current ratings. Furthermore, sentiment analysis can be used on language of reviews and matched with the star reviews to provide more information on a business?s chances of success. Additionally, information about the current state of a business and likely future states could be provided to businesses to permit them to change their current practices and focus to increase their chances of success.

@@

A previous winner, Hood et al., published a paper on assessing a business?s current rating/review state and future expectations with respect to review numbers on Yelp. The bulk of their work centered around the creation of additional features for the data (e.g. number of similar businesses within 1km, features of subsets of reviews such as average rating and number of unique users). From there they utilized various feature reduction and clustering techniques to build a prediction model. They were able to identify features that offered greater accuracy in prediction of interest in the business (calculated as the predicted number of reviews within 6 months after a target date) as well as a create model for prediction that outperformed a basic linear regression. Our approach will be similar in the application of feature reduction to reduce the dimensions of that data and clustering to find cities with many businesses, as well as to group similar businesses together. However, we do not plan to create additional features like the number of nearby similar businesses.

@@

Huang et al. (2013) utilized probabilistic models to discover underlying subtopics within Yelp reviews. This offered a framework to categorize the reviews based on review text. From here, they were able to analyze the prevalence of various star ratings within each category. In this way, they could advise business owners as to the common trends among good and bad reviews that may not be apparent without considering the review text. McAuley and Leskovec built on this idea by combining such latent review topics with hidden rating driving factors, allowing them construct a rating prediction model.

@@

The dataset has many optional attributes. There is an attribute simply called 'attributes' that is a shopping list of optional fields for each business. This part of the dataset is inconsistent and the same fields cannot be relied upon to be present for each business. For instance, one of these attributes is 'Bike Parking.' This attribute is either true, false, or nonexistent. In the cases when it is nonexistent, analysis based upon this attribute is obviously unreliable. Most of these miscellaneous attributes are present for much less than a majority of the data, so any mining based upon them would have to first drastically trim the data available for analysis by eliminating more than half the data. For this reason, we have chosen to eliminate the miscellaneous category rather than most of our data.

\quad In addition, the data is not only restaurants but many kinds of businesses. A datum associated with each business attempts to classify the business as a member of a category, such as ?restaurants,? ?beaches,? or ?shopping.? However some categories are extremely broad and the same type of business can be described in different ways. One example is that a bakery may be classified as belonging to ?bakeries,? but also generically as ?food.? Conversely, many businesses sell food, but the generic category ?food? will not be in their list of categories and instead the business will be labelled with more specific categories like ?Italian food,? diminishing the usefulness of these categories.

\quad The dataset provided by the Yelp organizers is fairly clean already, and the majority of the data cleaning process will be involved in handling missing data. It is possible to fill in these missing values with default values like ?false?, or ?none? but it can be dangerous to make assumptions about what businesses have and don?t have - for example, filling in a ?gluten-free? attribute of a business may influence our analysis of its reviews and, depending on exposure, maybe even mislead Yelp users as to what the business provides.

@@

The Yelp dataset comes from the Yelp Dataset Challenge. This is an ongoing challenge and this is the ninth round, so the challenge is very well-established by this point. The challenge offers \$5,000 to the winner and the final submission takes the form of an academic paper.

@@

\begin{itemize}
	\item{4.1 million reviews}
	\begin{verbatim}
	"review_id":(encrypted review id),
	"user_id":(encrypted user id),
	"business_id":(encrypted business id)"
	"stars":(star rating, rounded to half-stars),
	"date":(date, formatted: YYYY-MM-DD),
	"text":"review text",
	"useful":('useful' vote count),
	"funny":('funny' vote count),
	"cool": ('cool' vote count),
	"type": "review"
	\end{verbatim}
	\item{Information on 144,000 businesses in 11 cities from 4 countries (incl. the US, Germany, Canada, and the U.K. (with 1.1 million business attributes)}
	\begin{verbatim}
	"business_id":(encrypted business id),
	"name":"business name",
	"neighborhood": "hood name",
	"address":"full address",
	"city":"city",
	"state":"state"(if applicable),
	"postal code":"postal code",
	"latitude":latitude,
	"longitude":longitude,
	"stars":(star rating, rounded to half-stars),
	"review_count":(number of reviews),
	"is_open":0/1 (closed/open),
	"attributes":[(localized attribute tags)],
	"categories":[(localized category names)],
	"hours":[(hours strings)],
	"type": "business"
	\end{verbatim}
	\item{Over 1M users}
	\begin{verbatim}
	"user_id":(encrypted user id),
	"name":"first name",
	"review_count":(number of reviews),
	"yelping_since": (date, formatted: YYYY-MM-DD),
	"friends":[(encrypted user ids)],
	"useful":('useful' vote count sent by user),
	"funny":('funny' vote count sent by user),
	"cool": ('cool' vote count sent by user),
	"fans":"number of fans the user has",
	"elite":["an array of years the user was elite"],
	"average_stars":floating point average like 4.31,
	"compliment_type": (compliment count),
		...(same for each complement type)...
	"type":"user"
	\end{verbatim}
	\item{Over 200,000 user-submitted photographs, 947,000 tips, and aggregated check-in data for most businesses}
\end{itemize}

@@

\quad The tools and methods used for this project will no longer include a database for storing or integrating the data. Due to difficulties described in section 6 and relating to the unknown JSON variant used by the preparers of the Yelp dataset, storing the data in a database is not feasible at this time. Given that the data is already quite clean save for the format, with minimal replacement of missing values and some other changes necessary, it is possible for each group member to have a copy of the data stored locally that may be processed with our tools, with smaller subsets of data, results, and scripts shared via GitHub. The datasets will be merged locally, and while this does increase the amount of computational work necessary for each member?s machine, the dataset is still of a small enough size (less than 10 GB, excluding the photos) that this is practical. We will still use various Python libraries and modules for cleaning and preprocessing, and various other tools for transformation, mining, pattern evaluation, and knowledge presentation. The lack of a centralized database in removes the need for a number of Python libraries like Dora and FTFY ('fixes text for you') intended for data cleaning and preprocessing, but increases the necessity of group communication to reduce duplication of work and miscommunications. We currently use and will continue to use Slack, email, and in-person meetings to communicate and synchronize our efforts.

\quad Numpy, Scipy, and Pandas will be used to transform and mine the data, and other Python modules like Seaborn and Plotly (a Python API for D3) will be used for knowledge visualization and presentation. Rather than relying on tools like Weka, the Text Analytics API from Microsoft Cognitive Services, or the Facebook fastText pre-trained word vectors, we will instead use Python?s Natural Language Toolkit for sentiment analysis and other language processing.

@@

\quad Evaluating our predictions as to whether a business success will hinge on comparing the star rating (from one to five stars) and our predicted star rating based on the other attributes in the review, such as day of the week and month, season, or location. This could be done by training an ML algorithm like Kernel Ridge regression on 80\% of the data and testing it on the remaining 20\%, and then adjusting parameters to minimize the error between the actual star ratings and our predicted star ratings for the testing data.

@@

There have been several major changes to the project tools and methods since the project proposal. Initially we planned on using a relational database to store, integrate, and access the data, but but we encountered many difficulties in just uploading the data. The inconsistencies between the user and business data made it exceedingly difficult to successfully import the data en masse into the database, so instead we decided to import the data straight into Weka. However, the format of the JSON file also did not allow a successful import into Weka.

@@

After importing the CSV version of the data, we tested some basic visualization tools from RapidMiner. For instance, in Figure 1 we created a histogram of the average star for a business versus the state in which it is located. Basic graphs like these can help refine our questions. For example, why does Illinois have significantly lower average stars than other states or provinces, even those that border it?

\quad The most pervasive issue with our business data is that many entries are missing data, most notably the miscellaneous 'attributes' and address information. Since these values are managed by the account owner, they vary widely across the businesses. However, some of these data are redundant given the other fields, such as detailed latitude-longitude data for each business.

\begin{figure}[h]
\includegraphics[width=0.75\textwidth, center]{b_rating_by_state}
\end{figure}

RapidMiner can also produce 3D scatterplots which is useful because we plan to use latitude and longitude to locate businesses, since some of the business address data is missing and cannot (and should not) be replaced with default values or other common values used to replace missing data. Figure 2 depicts a graph of longitude vs. latitude vs. review count for businesses.

\begin{figure}[!h]
\includegraphics[width=0.75\textwidth, center]{3d_businesses_lat_long}
\end{figure}

In this instance the visualization makes it clear that reviews are more frequent in populated cities as we expect. However, it also shows visually the outliers. There are various businesses that have only a few reviews. In this context it is easier to identify outliers and also see that businesses concentrate in certain locations, as is expected given Yelp?s declaration that the dataset focuses on businesses in 11 cities. Therefore we can reduce the dataset to only the cities that meet the threshold of businesses or total review counts.

@@

One of the major goals of the project is to find recurrent problems and other issues with the businesses, and this was to be done by grouping reviews by business and then analyzing trends in the language used in the reviews over time. However, given that we were unable to find a suitable workaround for our difficulties with Weka, we used sentiment analysis tools from the Natural Language Toolkit (NLTK) to analyze the review language instead. NLTK is a language-processing and analysis platform that uses Python, and a good replacement for the language processing capabilities of Weka.

\quad Using a custom script (see GitHub\; nltk\_review\_subset\_process.py) to analyze the sentiment of the first 100,000 reviews (about 2.5\% of the available data), we discovered that the average sentiment score among all analyzed reviews was around 0.113, and there were 17295 negative reviews compared to 76968 positive reviews. The remaining 5737 reviews had composite sentiment scores within -0.2 < score < 0.2, so they were excluded from further analysis since they were neutral. This means that the average Yelp review (at least within the analyzed subset) is positive rather than negative or neutral in sentiment, perhaps indicating that Yelp users are more inclined to write reviews of businesses with which they had a good experience. Only around 2.5\% of the data is currently processed by the file, because processing the entire file is prohibitively expensive.

@@

The users dataset essentially provides a window into users' status, connections, and interaction within the Yelp ecosystem. The figure below shows a plot of the user data. We see that the user density grows around 3.25-4.25 average rating with some traces of a bell curve regarding the number of reviews.

\begin{figure}[!h]
\includegraphics[width=0.45\textwidth, center]{u_reviews_vs_stars}
\end{figure}

\quad One of the most significant yet hardest to process attributes is the 'friends' list. This attribute provides an opportunity to model our data as a graph. Such a network could provide very relevant clustering through sparseness measures and cutting in addition to accounting for user/business influence or finding deeper trends according to network connectivity. However, given the magnitude of the dataset, modeling the ?friends? of each user as a complete graph will be computationally difficult. To combat this, we could limit our model to the first n people the graph builder encounters only, since connections between groups of people are unpredictable.

\quad
\section{Next Steps}

@@

\quad Currently, the review dataset is too large to process in detail. In the future, we will continue taking subset samples of a more reasonable size and processing those instead, as well as increasing the complexity of our analysis. Combining the review dataset with the business dataset (using the 'business\_id' field as a key) will enable us to produce maps of where the largest number of reviews with negative and positive sentiment are, as well as seeing where users leave the most strongly-worded reviews. We hope that additional advanced sentiment analysis will be useful in determining underlying factors that influence a business's success, as well as in identifying recurrent problems.
